---
title: "Guided Exercise 3: Box Office Predictions"
Author: Bobby Kent, Riley Black, Benjamin Fahrenkrug, Cecily Shi
output: html_document
editor_options: 
  chunk_output_type: console
---

## Your First Kaggle Competition

[Kaggle](<https://www.kaggle.com/>) is a data science community that hosts competitions. We're going to host our very own Kaggle style competition for this assignment. 

### Outcome

Your goal is to predict _log_ inflation-adjusted box office take, `box_office_a`, from a dataset containing information on movies released since 2000.  

### Criteria

We will use log root mean square error for this competition. This is the same rmse we've been using in class. 

### Prize

The prize will be the best one we can possible give you: the respect of your peers. 

### How it works

We will give you two datasets. The first `movie_train.Rds` contains complete information on 4,338 movies. The second, `movie_test.Rds` will have data on 482 movies, but will not contain any informatin on the outcomes. You need to train a model on the `movie_train` data and then generate predictions from the `movie_test` data. We will compare your predictions with the actual box office take of the movies in the testing dataset, then report your rmse back to you. 

You may NOT use any other data sources to  supplement what we give you. Please stick with this rule, this is just for fun in any case. 

### Submission

You need to send us a .csv file (use `write_csv` to export it) containing only two elements. The first column should be `imdbid`, the second column-- predicted box office take--- should be `pred`. We will calculate the rmse for you and report your score on the dates listed below. Please email your results to the class email address. 

### Reporting Scores

We will report your score to you on the following dates, provide we have a submission at that time:

November 1, November 3, November 5, November 7


## Codebook for Movie Dataset

|        *Variable*        |             *Description*        |
|--------------|---------------------------------------------|
| imdbid       | id                                          |
| title        | title                                       |
| year         | year                                        |
| rated        | rated                                       |
| released     | Date of release                             |
| runtime      | Running time in minutes                     |
| genre1       | Genre                                       |
| genre2       | Genre                                       |
| genre3       | Genre                                       |
| director1    | Director                                    |
| director2    | Director                                    |
| director3    | Director                                    |
| writer1      | Writer                                      |
| writer2      | Writer                                      |
| writer3      | Writer                                      |
| actor1       | Actor                                       |
| actor2       | Actor                                       |
| actor3       | Actor                                       |
| plot         | Plot Summary                                |
| language1    | Language                                    |
| language2    | Language                                    |
| language3    | Language                                    |
| country1     | Country                                     |
| country2     | Country                                     |
| country3     | Country                                     |
| awards       | Text descriptiono of nominations and awards |
| metascore    | Meta Score from Rotten Tomatoes             |
| imdbrating   | IMDB Rating (1:10)                          |
| imdbvotes    | Number of Votes on IMDB                     |
| production1  | Production Company                          |
| production2  | Production Company                          |
| production3  | Production Company                          |
| boxoffice_a  | Inflation adjusted boxoffice take           |
| oscar_winner | Did movie win any Oscar awards (1=yes,0=no) |
|              |                                             |

# Data Wrangling

## Data Import

```{r}
library(tidyverse)
library(tidymodels)
library(knitr)
library(plotly)
library(modelr)
```

```{r}
#Import Data
movie_test <-readRDS("movie_test.Rds")
movie_train <-readRDS("movie_train.Rds")
```
## Category Combination and Creation of Category Binary Variables

### Combine 3 Genre Columns and Create Binary Var Columns for each Genre Type

```{r}
genre_table <- movie_train %>% 
  group_by(genre1) %>% # Group by types of genres seen in data set
  count() %>%
  arrange(by=-n) # Arrange from least to most common
glimpse(genre_table)

total_genres <- colSums(genre_table[,-1])

genre_table_above_threshold <- genre_table %>%
  filter(n/total_genres >= 0.005)
genre_list <- genre_table_above_threshold[1] # Generate list (1 col) of genre types
```

```{r}
for (i in 1:nrow(genre_list)) { # Iterate through our genre list
  genre <- as.character(genre_list[i,]) # Save current genre as chr type
  movie_train <- movie_train %>% # Create new column with name equal to the current genre and check to see if any of the three genre columns are that kind. If they are we will set this value to 1. Otherwise we will set it to 0.
  mutate({{genre}} := if_else(((!is.na(genre1) & genre1== {{genre}}) | (!is.na(genre2) & genre2== {{genre}}) | (!is.na(genre3) & genre3== {{genre}})), 1, 0))
}
```

### Combine 3 Director Columns and Create Binary Var Columns for each Director

```{r}
director_table <- movie_train %>% 
  group_by(director1) %>% # Group by types of directors seen in data set
  count() %>%
  arrange(by=-n) # Arrange from least to most common
glimpse(director_table)

total_directors <- colSums(director_table[,-1])

director_table_above_threshold <- director_table %>%
  filter(n/total_directors >= 0.002)
director_list <- director_table_above_threshold[1] # Generate list (1 col) of director types
```

```{r}
for (i in 1:nrow(director_list)) { # Iterate through our director list
  director <- as.character(director_list[i,]) # Save current director as chr type
  movie_train <- movie_train %>% # Create new column with name equal to the current director and check to see if any of the three director columns are that kind. If they are we will set this value to 1. Otherwise we will set it to 0.
  mutate({{director}} := if_else(((!is.na(director1) & director1== {{director}}) | (!is.na(director2) & director2== {{director}}) | (!is.na(director3) & director3== {{director}})), 1, 0))
}
```

### Combine 3 Writer Columns and Create Binary Var Columns for each Writer

```{r}
writer_table <- movie_train %>% 
  group_by(writer1) %>% # Group by types of writers seen in data set
  count() %>%
  arrange(by=-n) # Arrange from least to most common
glimpse(writer_table)

total_writers <- colSums(writer_table[,-1])

writer_table_above_threshold <- writer_table %>%
  filter(n/total_writers >= 0.002)
writer_list <- writer_table_above_threshold[1] # Generate list (1 col) of writer types
```

```{r}
for (i in 1:nrow(writer_list)) { # Iterate through our writer list
  writer <- as.character(writer_list[i,]) # Save current writer as chr type
  movie_train <- movie_train %>% # Create new column with name equal to the current writer and check to see if any of the three writer columns are that kind. If they are we will set this value to 1. Otherwise we will set it to 0.
  mutate({{writer}} := if_else(((!is.na(writer1) & writer1== {{writer}}) | (!is.na(writer2) & writer2== {{writer}}) | (!is.na(writer3) & writer3== {{writer}})), 1, 0))
}
```

### Combine 3 Actor Columns and Create Binary Var Columns for each Actor

```{r}
actor_table <- movie_train %>% 
  group_by(actor1) %>% # Group by types of actors seen in data set
  count() %>%
  arrange(by=-n) # Arrange from least to most common
glimpse(actor_table)

total_actors <- colSums(actor_table[,-1])

actor_table_above_threshold <- actor_table %>%
  filter(n/total_actors >= 0.002)
actor_list <- actor_table_above_threshold[1] # Generate list (1 col) of actor types
```

```{r}
for (i in 1:nrow(actor_list)) { # Iterate through our actor list
  actor <- as.character(actor_list[i,]) # Save current actor as chr type
  movie_train <- movie_train %>% # Create new column with name equal to the current actor and check to see if any of the three actor columns are that kind. If they are we will set this value to 1. Otherwise we will set it to 0.
  mutate({{actor}} := if_else(((!is.na(actor1) & actor1== {{actor}}) | (!is.na(actor2) & actor2== {{actor}}) | (!is.na(actor3) & actor3== {{actor}})), 1, 0))
}
```

### Combine 3 Language Columns and Create Binary Var Columns for each Language Type

```{r}
language_table <- movie_train %>% 
  group_by(language1) %>% # Group by types of languages seen in data set
  count() %>%
  arrange(by=-n) # Arrange from least to most common
glimpse(language_table)

total_languages <- colSums(language_table[,-1])

language_table_above_threshold <- language_table %>%
  filter(n/total_languages >= 0.005)
language_list <- language_table_above_threshold[1] # Generate list (1 col) of language types
```

```{r}
for (i in 1:nrow(language_list)) { # Iterate through our language list
  language <- as.character(language_list[i,]) # Save current language as chr type
  movie_train <- movie_train %>% # Create new column with name equal to the current language and check to see if any of the three language columns are that kind. If they are we will set this value to 1. Otherwise we will set it to 0.
  mutate({{language}} := if_else(((!is.na(language1) & language1== {{language}}) | (!is.na(language2) & language2== {{language}}) | (!is.na(language3) & language3== {{language}})), 1, 0))
}
```

### Combine 3 Country Columns and Create Binary Var Columns for each Country Type

```{r}
country_table <- movie_train %>% 
  group_by(country1) %>% # Group by types of countrys seen in data set
  count() %>%
  arrange(by=-n) # Arrange from least to most common
glimpse(country_table)

total_countrys <- colSums(country_table[,-1])

country_table_above_threshold <- country_table %>%
  filter(n/total_countrys >= 0.05)
country_list <- country_table_above_threshold[1] # Generate list (1 col) of country types
```

```{r}
for (i in 1:nrow(country_list)) { # Iterate through our country list
  country <- as.character(country_list[i,]) # Save current country as chr type
  movie_train <- movie_train %>% # Create new column with name equal to the current country and check to see if any of the three country columns are that kind. If they are we will set this value to 1. Otherwise we will set it to 0.
  mutate({{country}} := if_else(((!is.na(country1) & country1== {{country}}) | (!is.na(country2) & country2== {{country}}) | (!is.na(country3) & country3== {{country}})), 1, 0))
}
```

### Combine 3 Production Columns and Create Binary Var Columns for each Production Type

```{r}
production_table <- movie_train %>% 
  group_by(production1) %>% # Group by types of productions seen in data set
  count() %>%
  arrange(by=-n) # Arrange from least to most common
glimpse(production_table)

total_productions <- colSums(production_table[,-1])

production_table_above_threshold <- production_table %>%
  filter(n/total_productions >= 0.05)
production_list <- production_table_above_threshold[1] # Generate list (1 col) of production types
```

```{r}
for (i in 1:nrow(production_list)) { # Iterate through our production list
  production <- as.character(production_list[i,]) # Save current production as chr type
  movie_train <- movie_train %>% # Create new column with name equal to the current production and check to see if any of the three production columns are that kind. If they are we will set this value to 1. Otherwise we will set it to 0.
  mutate({{production}} := if_else(((!is.na(production1) & production1== {{production}}) | (!is.na(production2) & production2== {{production}}) | (!is.na(production3) & production3== {{production}})), 1, 0))
}
```

Now, we have completed the creation of binary categories for each of the above threshold genres, directors, writers, actors, languages, countries, and productions. This means that we can now use these variables in our Lasso analysis below along with the other relevant variables in our data set.